Refining from the BON High-level Specification Language to the VHDL-AMS Hardware Description Language

### 2009-07-29 papers/BON_VHDL/paper.tex
paper progress

\newcommand{\tablesize}{\footnotesize}
\newcommand{\eg}{e.g.,\xspace}
\newcommand{\ie}{i.e.,\xspace}
\newcommand{\etc}{etc.\xspace}
\newcommand{\myhref}[2]{\ifpdf\href{#1}{#2}\else\htmladdnormallinkfoot{#2}{#1}\fi}
%\newcommand{\myhref}[2]{\emph{#2}}
\newcommand{\todo}{\textbf{TODO:}}
\newcommand{\ST}{\emph{SenseTile}\xspace}
\newcommand{\STs}{\emph{SenseTiles}\xspace}
\newcommand{\STS}{\emph{SenseTile Simulator}\xspace}
\newcommand{\datastore}{\STs Scientific Datastore\xspace}
\newcommand{\computefarm}{The \STs Scientific Compute Farm\xspace}
\newcommand{\computefarmlong}{UCD CASL \STs Software and Data Compute Server Farm\xspace}
\newcommand{\sensorfarm}{The \STs\xspace}
\newcommand{\Sc}{\emph{system\_chart}\xspace}
\newcommand{\Clc}{\emph{cluster\_chart}\xspace}
\newcommand{\Csc}{\emph{class\_chart}\xspace}
\newcommand{\Ft}{\emph{feature}\xspace}
\newcommand{\ent}{\emph{entity}\xspace}
\newcommand{\arch}{\emph{architecture}\xspace}
\newcommand{\archs}{\emph{architectures}\xspace}
%---------------------------------------------------------------------
% New commands, macros, \etc
%---------------------------------------------------------------------

@@

The authors will illustrate in this paper how BON can represent a high 
level analogue mapping and will further illustrate a procedure for 
transformation from BON to VHDL-AMS. As an example, 
we will give practical aspect of refinement from BON to VHDL-AMS,
applied on A/D convertor, in order to produce formal specification 
and software simulation.
%=====================================================================

\section{VHDL-AMS}
VHDL-AMS is the IEEE standard modeling language (standard 1076.1) created 
to provide compatibility and capability in an open language for modern analog, 
mixed-signal, multi-domain designs.
\section{Choice of BON as the modeling tool}
%Brief description and motivation.
The Business Object Notation (BON) is an analysis and design notation 
for object-oriented systems  which emphasizes \emph{seamlessness}, 
\emph{reversibility} and software \emph{contracting}. Let us define 
these principles precisely and point out why the principles are 
important in VHDL-AMS simulation.
%\todo{dragan 25/07/2009 - Figure out relation: Seamlessness and VHDL-AMS}
\begin{itemize}
\item \emph{Seamlessness} is the use of a continuous process 
throughout the software lifecycle.
%\todo{dragan 25/07/2009 - Figure out relation: Reversibility and VHDL-AMS. Why is Reversibility important in VHDL design.}
\item \emph{Reversibility} is the support for both forward and 
backward development process: from analysis to design 
and implementation, and back.
In VHDL-AMS, modelling represents an continious top-down-bottom-up
process. Top-down approach serves for design while bottom-up
serves for verification. The possibility of seamlessly translating 
changes made during a certain development phase back into earlier phases,
facilitates a stable and predictable environment and ensures the various 
models consistent with the code.
\item \emph{Contracting} is the precise definition, for each software element, 
of the associated semantic properties. BON use a full assertion mechanism, 
allowing analysts to specify not only the structure of a system but also its
semantics ( constraints and invariants of the expected results ).
\end{itemize}

@@

\subsection{Mapping between BON and VHDL-AMS}
%\todo{dragan 27/07/2009 - Figure out : Mapping from BON  to entity/architecture structure in VHDL-ASM.}
\begin{itemize}
\item \Sc: can be represented with high abstract level of the topmost 
entity in a design hierarchy.
\item \Clc: to be described.
\item \Csc: can be identify with the primary unit(\ent) and secondary 
unit (\arch) structure in VHDL-AMS.
The \Csc's name corresponds to the name of the entity in VHDL-AMS.
\item \Ft: The feature section of the \Csc, where the mathematical 
transfer function of the component is entered, holds the same information 
as the \arch section of the VHDL-AMS.
\end{itemize}
An VHDL-AMS component are described using an \ent-\arch structure. 
The component model contains an \ent and one or more \archs. 
The \ent specifies the interface of the model to the outside world. 
In other words, \ent introduces a name, list of input/output ports 
and the definition of the generic parameters. The textual 
presentation of \ent is shown in Figure.1. 
\input{adc-entity}

>>>>

\newcommand{\tablesize}{\footnotesize}
\newcommand{\eg}{e.g.,\xspace}
\newcommand{\ie}{i.e.,\xspace}
\newcommand{\etc}{etc.\xspace}
\newcommand{\myhref}[2]{\ifpdf\href{#1}{#2}\else\htmladdnormallinkfoot{#2}{#1}\fi}
%\newcommand{\myhref}[2]{\emph{#2}}
\newcommand{\todo}{\textbf{TODO:}}
\newcommand{\ST}{\emph{SenseTile}\xspace}
\newcommand{\STs}{\emph{SenseTiles}\xspace}
\newcommand{\STS}{\emph{SenseTile Simulator}\xspace}
\newcommand{\datastore}{\STs Scientific Datastore\xspace}
\newcommand{\computefarm}{The \STs Scientific Compute Farm\xspace}
\newcommand{\computefarmlong}{UCD CASL \STs Software and Data Compute Server Farm\xspace}
\newcommand{\sensorfarm}{The \STs\xspace}
\newcommand{\Sc}{\emph{system\_chart}\xspace}
\newcommand{\Clc}{\emph{cluster\_chart}\xspace}
\newcommand{\Csc}{\emph{class\_chart}\xspace}
\newcommand{\Ft}{\emph{feature}\xspace}
\newcommand{\ent}{\emph{entity}\xspace}
\newcommand{\arch}{\emph{architecture}\xspace}
\newcommand{\archs}{\emph{architectures}\xspace}
\newcommand{\duc}{\emph{design units collection}\xspace}
\newcommand{\ducs}{\emph{design units collections}\xspace}
\newcommand{\du}{\emph{design unit}\xspace}
%---------------------------------------------------------------------
% New commands, macros, \etc
%---------------------------------------------------------------------

@@

The authors will illustrate in this paper how BON can represent a high 
level analogue mapping and will further illustrate a procedure for 
transformation from BON to VHDL-AMS. As an extremly simple example, 
we will give practical aspect of refinement from BON to VHDL-AMS,
applied on A/D convertor, in order to produce formal specification 
and software simulation.
%=====================================================================

\section{VHDL-AMS}
IEEE VHDL-AMS 1076.1 is the industry standard mixed-signal 
high-level description language for electronic and multi-domain 
systems. VHDL-AMS, as an superset of VHDL, supports modeling and 
simulation of analog and mixed-signal systems. The main contribution 
of VHDL-AMS is ability to describe the behavior of complex system 
trought differential and algebraic equations (DAEs). For this purpose, 
VHDL-AMS introduces the following capabilities and therms:

\begin{itemize}
\item 
\emph{Terminals} represents phisical connection points or circuit nodes 
of a network connectivity. Terminals are declared to be of various 
\emph{natures}, which represent different energy domains of a system.
\item
The \emph{simultaneous statements} creates one or more characteristic 
equations that are used by the "analog solver" to model continuously 
valued behaviour.  
\item
The \emph{quantities} denote a waveform or a time series of values.
The behaviour of the \emph{quantities} is given by the \emph{simultaneous
 statements}. They take values as a result of solving the set of simultaneous 
ordinary differential and algebraic equations (DAEs). To simplify, 
quantities can be understood as analog unknowns of (DAEs) that specify 
analog behavior. 
\item
The \emph{break statement} is an construct which allows to 
initialize the DAE system at the beginning of a transient 
simulation and to mark discontinuities during such a transient 
simulation.

\end{itemize}

\section{Choice of BON as the modeling tool}
%Brief description and motivation.
The Business Object Notation (BON) is an analysis and design notation 
for object-oriented systems  which emphasizes \emph{seamlessness}, 
\emph{reversibility} and software \emph{contracting}. Let us define 
these principles precisely and point out why the principles are 
important in VHDL-AMS simulation.
%\todo{dragan 25/07/2009 - Figure out relation: Seamlessness and VHDL-AMS}
\begin{itemize}
\item \emph{Seamlessness} is the use of a continuous process 
throughout the software lifecycle.
%\todo{dragan 25/07/2009 - Figure out relation: Reversibility and VHDL-AMS. Why is Reversibility important in VHDL design.}
\item \emph{Reversibility} is the support for both forward and 
backward development process: from analysis to design 
and implementation, and back.
In VHDL-AMS, modelling represents an continious top-down-bottom-up
process. Top-down approach serves for design while bottom-up
serves for verification. The possibility of seamlessly translating 
changes made during a certain development phase back into earlier phases,
facilitates a stable and predictable environment and ensures the various 
models consistent with the code.
\item \emph{Contracting} is the precise definition, for each software element, 
of the associated semantic properties. BON use a full assertion mechanism, 
allowing analysts to specify not only the structure of a system but also its
semantics ( constraints and invariants of the expected results ).
\end{itemize}

@@

\subsection{Mapping between BON and VHDL-AMS}
%\todo{dragan 27/07/2009 - Figure out : Mapping from BON  to entity/architecture structure in VHDL-ASM.}
\begin{itemize}
\item \Sc: VHDL-AMS allows hierarchical structure modeling, 
meaning that subsistems ( \ducs ), can be developed independently.
As an top-level description (informal/formal), \Sc connect all the \ducs
together to a complete design. 
\item \Clc: An  subsistem which represent \duc.
\item \Csc: describes \du. It can be identify with the primary unit (\ent) and secondary 
unit (\arch) structure in VHDL-AMS.
The \Csc's name corresponds to the name of the entity in VHDL-AMS.
\item \Ft: The feature section of the \Csc, where the mathematical 
transfer function of the component is entered, holds the same information 
as the \arch section of the VHDL-AMS.
\end{itemize}
An VHDL-AMS component are described using an \ent-\arch structure. 
The component model contains an \ent and one or more \archs. 
The \ent specifies the interface of the model to the outside world. 
In other words, \ent introduces a name, list of input/output ports 
and the definition of the generic parameters. The textual 
presentation of \ent is shown in Figure.1. 
\input{adc-entity}
